---
title: "Clustering Examples (RSWN)"
author: "Tom Helmuth"
date: "March 3, 2015"
output: 
  html_document:
    toc: true
---

# Introduction: Replace Space With Newline problem

This document will make clustering plots for data from a range of settings and results.

## Setup code

```{r, message=FALSE}
library('ggplot2')
library('cluster')
library('apcluster')

setwd("~/Documents/R/Clustering/lexicase-clusturing-analysis/R_notebooks")

source('../scripts/clustering.R')
```

## Run These Once To Get Data Into CSVs

These transform the original data files into error files, and transform the error data into data frames we can use.

Lexicase, Run 1, a success
```{r}
# transform_data_file_into_error_file("data/RSWN/lexicase/data1.csv")
# data_frame1 = make_frame_from_errors_file(
#   "data/RSWN/lexicase/errors_data1.csv", 1, 
#   "replace-space-with-newline", "lexicase", 20, elitize_generation_data)
# write.csv(data_frame1, "data/RSWN/lexicase/error_counts_and_div1.csv")
```

Lexicase, Run 6, a success
```{r}
# transform_data_file_into_error_file("data/RSWN/lexicase/data6.csv")
# data_frame6 = make_frame_from_errors_file(
#   "data/RSWN/lexicase/errors_data6.csv", 6, 
#   "replace-space-with-newline", "lexicase", 20, elitize_generation_data)
# write.csv(data_frame6, "data/RSWN/lexicase/error_counts_and_div6.csv")
```

Lexicase, Run 2, a failure
```{r}
# transform_data_file_into_error_file("../data/RSWN/lexicase/data2.csv")
# data_frame2 = make_frame_from_errors_file(
#   "../data/RSWN/lexicase/errors_data2.csv", 2, 
#   "replace-space-with-newline", "lexicase", 20, elitize_generation_data)
# write.csv(data_frame2, "../data/RSWN/lexicase/error_counts_and_div2.csv")
```

Lexicase, Run 3, a failure
```{r}
# transform_data_file_into_error_file("../data/RSWN/lexicase/data3.csv")
# data_frame3 = make_frame_from_errors_file(
#   "../data/RSWN/lexicase/errors_data3.csv", 3, 
#   "replace-space-with-newline", "lexicase", 20, elitize_generation_data)
# write.csv(data_frame3, "../data/RSWN/lexicase/error_counts_and_div3.csv")
```

# Lexicase, Success

```{r}
data_frame1 = read.csv("../data/RSWN/lexicase/error_counts_and_div1.csv")
data_frame6 = read.csv("../data/RSWN/lexicase/error_counts_and_div6.csv")

lexicase_success <- rbind(data_frame1, data_frame6)
```

## Clusters
```{r}
ggplot(lexicase_success, aes(x=generation)) + 
  geom_line(aes(y=cluster.count, color=interaction("Run", run.num))) + 
  labs(y="Number of Clusters", color="")
```

## Diversity
```{r}
ggplot(lexicase_success, aes(x=generation)) + 
  geom_line(aes(y=error.diversity, color=interaction("Run", run.num))) + 
  labs(y="Error Diversity", color="")
```

# Lexicase, Failure

```{r}
data_frame2 = read.csv("../data/RSWN/lexicase/error_counts_and_div2.csv")
data_frame3 = read.csv("../data/RSWN/lexicase/error_counts_and_div3.csv")

lexicase_failure <- rbind(data_frame2, data_frame3)
```

## Clusters
```{r}
ggplot(lexicase_failure, aes(x=generation)) + 
  geom_line(aes(y=cluster.count, color=interaction("Run", run.num))) + 
  labs(y="Number of Clusters", color="")
```

## Diversity
```{r}
ggplot(lexicase_failure, aes(x=generation)) + 
  geom_line(aes(y=error.diversity, color=interaction("Run", run.num))) + 
  labs(y="Error Diversity", color="")
```

