---
title: "Clustering Lexicase-Tournament Selection"
author: "Tom Helmuth"
date: "June 1, 2015"
output: 
  html_document:
    toc: true
---

# Introduction

The goal of this document is to look at the clustering and diversity data from the lexicase-touranment runs, and compare them with lexicase and tournament individually.

## Setup Code

```{r, message=FALSE}
library('ggplot2')
library('cluster')
library('Hmisc')
library("gridExtra")
library("plyr")

setwd("~/Documents/R/Clustering/lexicase-clusturing-analysis/R_notebooks")

source('../scripts/clustering.R')
```

## Import Data

### Replace Space With Newline

```{r}
data_rswn_lexicase = import_from_error_clustering_and_div("../data/RSWN/lexicase/clustering/")
data_rswn_tourney = import_from_error_clustering_and_div("../data/RSWN/tourney/clustering/")
data_rswn_ifs = import_from_error_clustering_and_div("../data/RSWN/ifs/clustering/")
data_rswn_ltk = import_from_error_clustering_and_div("../data/RSWN/lexicase-tourney-keep/clustering/")
data_rswn_ltr = import_from_error_clustering_and_div("../data/RSWN/lexicase-tourney-remove/clustering/")
data_rswn = rbind(data_rswn_lexicase, data_rswn_tourney, data_rswn_ifs, data_rswn_ltk, data_rswn_ltr)
data_rswn$succeeded = factor(data_rswn$succeeded, levels = c(TRUE, FALSE))
data_rswn$treatment = factor(data_rswn$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```

### Syllables

```{r}
data_syllables_lexicase = import_from_error_clustering_and_div("../data/syllables/lexicase/clustering/")
data_syllables_tourney = import_from_error_clustering_and_div("../data/syllables/tourney/clustering/")
data_syllables_ifs = import_from_error_clustering_and_div("../data/syllables/ifs/clustering/")
data_syllables_ltk = import_from_error_clustering_and_div("../data/syllables/lexicase-tourney-keep/clustering/")
data_syllables_ltr = import_from_error_clustering_and_div("../data/syllables/lexicase-tourney-remove/clustering/")
data_syllables = rbind(data_syllables_lexicase, data_syllables_tourney, data_syllables_ifs, data_syllables_ltk, data_syllables_ltr)
data_syllables$succeeded = factor(data_syllables$succeeded, levels = c(TRUE, FALSE))
data_syllables$treatment = factor(data_syllables$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```

### String Lengths Backwards

```{r}
data_string_lengths_backwards_lexicase = import_from_error_clustering_and_div("../data/string-lengths-backwards/lexicase/clustering/")
data_string_lengths_backwards_tourney = import_from_error_clustering_and_div("../data/string-lengths-backwards/tourney/clustering/")
data_string_lengths_backwards_ifs = import_from_error_clustering_and_div("../data/string-lengths-backwards/ifs/clustering/")
data_string_lengths_backwards_ltk = import_from_error_clustering_and_div("../data/string-lengths-backwards/lexicase-tourney-keep/clustering/")
data_string_lengths_backwards_ltr = import_from_error_clustering_and_div("../data/string-lengths-backwards/lexicase-tourney-remove/clustering/")
data_string_lengths_backwards = rbind(data_string_lengths_backwards_lexicase, data_string_lengths_backwards_tourney, data_string_lengths_backwards_ifs, data_string_lengths_backwards_ltk, data_string_lengths_backwards_ltr)
data_string_lengths_backwards$succeeded = factor(data_string_lengths_backwards$succeeded, levels = c(TRUE, FALSE))
data_string_lengths_backwards$treatment = factor(data_string_lengths_backwards$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```

### Negative To Zero

```{r}
data_negative_to_zero_lexicase = import_from_error_clustering_and_div("../data/negative-to-zero/lexicase/clustering/")
data_negative_to_zero_tourney = import_from_error_clustering_and_div("../data/negative-to-zero/tourney/clustering/")
data_negative_to_zero_ifs = import_from_error_clustering_and_div("../data/negative-to-zero/ifs/clustering/")
data_negative_to_zero_ltk = import_from_error_clustering_and_div("../data/negative-to-zero/lexicase-tourney-keep/clustering/")
data_negative_to_zero_ltr = import_from_error_clustering_and_div("../data/negative-to-zero/lexicase-tourney-remove/clustering/")
data_negative_to_zero = rbind(data_negative_to_zero_lexicase, data_negative_to_zero_tourney, data_negative_to_zero_ifs, data_negative_to_zero_ltk, data_negative_to_zero_ltr)
data_negative_to_zero$succeeded = factor(data_negative_to_zero$succeeded, levels = c(TRUE, FALSE))
data_negative_to_zero$treatment = factor(data_negative_to_zero$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```

### Doubles Letters

```{r}
data_double_letters_lexicase = import_from_error_clustering_and_div("../data/double-letters/lexicase/clustering/")
data_double_letters_tourney = import_from_error_clustering_and_div("../data/double-letters/tourney/clustering/")
data_double_letters_ifs = import_from_error_clustering_and_div("../data/double-letters/ifs/clustering/")
data_double_letters_ltk = import_from_error_clustering_and_div("../data/double-letters/lexicase-tourney-keep/clustering/")
data_double_letters_ltr = import_from_error_clustering_and_div("../data/double-letters/lexicase-tourney-remove/clustering/")
data_double_letters = rbind(data_double_letters_lexicase, data_double_letters_tourney, data_double_letters_ifs, data_double_letters_ltk, data_double_letters_ltr)
data_double_letters$succeeded = factor(data_double_letters$succeeded, levels = c(TRUE, FALSE))
data_double_letters$treatment = factor(data_double_letters$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```

### Count Odds

```{r}
data_count_odds_lexicase = import_from_error_clustering_and_div("../data/count-odds/lexicase/clustering/")
data_count_odds_tourney = import_from_error_clustering_and_div("../data/count-odds/tourney/clustering/")
data_count_odds_ifs = import_from_error_clustering_and_div("../data/count-odds/ifs/clustering/")
data_count_odds_ltk = import_from_error_clustering_and_div("../data/count-odds/lexicase-tourney-keep/clustering/")
data_count_odds_ltr = import_from_error_clustering_and_div("../data/count-odds/lexicase-tourney-remove/clustering/")
data_count_odds = rbind(data_count_odds_lexicase, data_count_odds_tourney, data_count_odds_ifs, data_count_odds_ltk, data_count_odds_ltr)
data_count_odds$succeeded = factor(data_count_odds$succeeded, levels = c(TRUE, FALSE))
data_count_odds$treatment = factor(data_count_odds$treatment, levels = c("lexicase", "tourney", "ifs", "lexicase-tourney-keep", "lexicase-tourney-remove"))
```


# Plot Medians and Quartiles

## RSWN

```{r}
source('../scripts/clustering.R')

add_generational_success_counts_plot(data_rswn, plot_diversity_medians_and_quartiles(data_rswn))

add_generational_success_counts_plot(data_rswn, plot_cluster_count_medians_and_quartiles(data_rswn))

#plot_cluster_count_medians_and_quartiles(data_rswn) + facet_grid(succeeded ~ ., labeller=label_both)
```

## Syllables

```{r}
add_generational_success_counts_plot(data_syllables, plot_diversity_medians_and_quartiles(data_syllables))

add_generational_success_counts_plot(data_syllables, plot_cluster_count_medians_and_quartiles(data_syllables))
```

## String Lengths Backwards

```{r}
add_generational_success_counts_plot(data_string_lengths_backwards, plot_diversity_medians_and_quartiles(data_string_lengths_backwards))

#plot_diversity_medians_and_quartiles(data_string_lengths_backwards) + facet_grid(succeeded ~ ., labeller=label_both)

add_generational_success_counts_plot(data_string_lengths_backwards, plot_cluster_count_medians_and_quartiles(data_string_lengths_backwards))

#plot_cluster_count_medians_and_quartiles(data_string_lengths_backwards) + facet_grid(succeeded ~ ., labeller=label_both)
```

## Negative To Zero

```{r}
add_generational_success_counts_plot(data_negative_to_zero, plot_diversity_medians_and_quartiles(data_negative_to_zero))

add_generational_success_counts_plot(data_negative_to_zero, plot_cluster_count_medians_and_quartiles(data_negative_to_zero))

#plot_cluster_count_medians_and_quartiles(data_negative_to_zero) + facet_grid(succeeded ~ ., labeller=label_both)
```

## Double Letters

```{r}
add_generational_success_counts_plot(data_double_letters, plot_diversity_medians_and_quartiles(data_double_letters))

add_generational_success_counts_plot(data_double_letters, plot_cluster_count_medians_and_quartiles(data_double_letters))

#plot_cluster_count_medians_and_quartiles(data_double_letters) + facet_grid(succeeded ~ ., labeller=label_both)
```

## Count Odds

```{r}
add_generational_success_counts_plot(data_count_odds, plot_diversity_medians_and_quartiles(data_count_odds))

add_generational_success_counts_plot(data_count_odds, plot_cluster_count_medians_and_quartiles(data_count_odds))
```