---
title: "Homology data"
author: "Nic McPhee and Thomas Helmuth"
date: "22 March 2015"
output: html_document
---

# **Important Note**

The `checksum` and `replace-space-with-newline` data here is from the same runs we used for the 
clustering and graph database work. The `negative-to-zero` and `string-length-backwards` data, however,
is from a _different_ collection of runs of those problems because the homology data wasn't output
for the clustering/graph DB runs for those problems.

# Load the libraries and data:

The reading/looping/binding code here assumes that this script is in our `R_notebooks` directory, and that that's the working directory. It then assumes the data is all in a `Homology_data` directory in `data`, with sub-directories for each problem, containing sub-directories for each treatment, each of which has a file `homology.csv` with the data for that problem and treatment.

I've commented out all that reading/looping/binding code because it's slow to re-build that combined data frame every time, and we'll just save and load the result.

```{r}
library("ggplot2")

# problem_names = c("checksum", "negative-to-zero", "replace-space-with-newline", "string-length-backwards")
# treatments = c("ifs", "lexicase", "tourney")
# 
# homology_data = data.frame()
# 
# for (p in problem_names) {
#   for (t in treatments) {
#     csv_file = paste0("../data/Homology_data/", p, "/", t, "/homology.csv")
#     data <- read.csv(csv_file)
#     data$problem = p
#     data$treatment = t
#     homology_data = rbind(homology_data, data)
#   }
# }
# 
# write.csv(homology_data, "../data/Homology_data/all_homology_data.csv", row.names=FALSE)

homology <- read.csv("../data/Homology_data/all_homology_data.csv")
```

# Plot just Run 6 of RSWN-lexicase

We can plot the 25%, 50%, and 75% data for a single run just fine:

```{r}
ggplot(subset(homology, problem=="replace-space-with-newline" & treatment=="lexicase" & Run==6), 
       aes(x=Generation, y=Homology, group=Quartile, color=factor(Quartile))) + 
  geom_line()
```

# Plot all the RSWN-lexicase runs

If we plot all the data with colors for the quartiles, it's just a muddy mess:

```{r}
ggplot(subset(homology, problem=="replace-space-with-newline" & treatment=="lexicase"), 
       aes(x=Generation, y=Homology, group=interaction(Run, Quartile), color=factor(Quartile))) + 
  geom_line(alpha=0.1)
```

If we split it out with factors, we find that they're really not all that different, although 
(as we'd expect) the numbers get higher as you move from 25% to 75%.

```{r}
ggplot(subset(homology, problem=="replace-space-with-newline" & treatment=="lexicase"), 
       aes(x=Generation, y=Homology, group=Run)) + 
  geom_line(alpha=0.1) + 
  facet_grid(. ~ Quartile)
```

Reading these plots is also complicated by (again) the fact that some runs end early, and it's hard to see when that's happening.

# Plotting medians for all three treatments

I tried several different plots, but everything tended towards the muddy mess, especially in black and 
white. So I just used facetting throughout.

## Checksum

```{r}
checksum_medians = subset(homology, problem=="checksum" & Quartile==0.5)

ggplot(checksum_medians, aes(x=Generation, y=Homology, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

The lexicase homology numbers are definitely quite a lot higher, and remarkabley tightly grouped 
throughout. (Lexicase isn't that tightly bunched in the other three problems.)
IFS and tournament look pretty much the same. These results are pretty similar to the diversity
plots, but don't tell us anything terribly useful about the lack of clustering.

It's interesting that in both IFS and Tourney you can see collapses in homology
(as nearly vertical lines) which probably represent moments where there's very strong convergence.
Those would be interesting moments to try to explore in the Graph DB.

## Negative-to-zero

```{r}
ntz_medians = subset(homology, problem=="negative-to-zero" & Quartile==0.5)

ggplot(ntz_medians, aes(x=Generation, y=Homology, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

This one's _really_ interesting because of the clear split between high and low homology runs for both 
IFS and Tourney, and for what appears to be some quantizing in both of those plots (horizontal lines with
multiple runs having the same median homology). It would be very interesting to see if those "high
homology" runs correspond to the few successes for IFS and Tourney.

## Replace space with newline

```{r}
rswn_medians = subset(homology, problem=="replace-space-with-newline" & Quartile==0.5)

ggplot(rswn_medians, aes(x=Generation, y=Homology, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

Again, both IFS and Tourney seem to split into "high" and "low" homology runs, with the high homology 
runs looking a lot more like lexicase than the low.

## String length backwards

```{r}
slb_medians = subset(homology, problem=="string-length-backwards" & Quartile==0.5)

ggplot(slb_medians, aes(x=Generation, y=Homology, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

This one's arguably the most interesting for lexicase, as it has the most variation through the run. 
Like `checksum` there appear to be some strong vertical drops in all three treatments that presumably 
represent some sort of diversity collapse.
