---
title: "Levenshtein distance data"
author: "Nic McPhee and Thomas Helmuth"
date: "22 March 2015"
output: 
  html_document:
    toc: true
---

# **Important Note**

The `checksum` and `replace-space-with-newline` data here is from the same runs we used for the 
clustering and graph database work. The `negative-to-zero` and `string-length-backwards` data, however,
is from a _different_ collection of runs of those problems because the Levenshtein distance data wasn't output
for the clustering/graph DB runs for those problems.

# Load the libraries and data:

The reading/looping/binding code here assumes that this script is in our `R_notebooks` directory, and that that's the working directory. It then assumes the data is all in a `Homology_data` directory in `data`, with sub-directories for each problem, containing sub-directories for each treatment, each of which has a file `homology.csv` with the data for that problem and treatment.

I've commented out all that reading/looping/binding code because it's slow to re-build that combined data frame every time, and we'll just save and load the result.

```{r}
library("ggplot2")
library(data.table)

# problem_names = c("checksum", "negative-to-zero", "replace-space-with-newline", "string-length-backwards")
# treatments = c("ifs", "lexicase", "tourney")
# 
# lev_dist_data = data.frame()
# 
# for (p in problem_names) {
#   for (t in treatments) {
#     csv_file = paste0("../data/Homology_data/", p, "/", t, "/homology.csv")
#     data <- read.csv(csv_file)
#     data$problem = p
#     data$treatment = t
#     lev_dist_data = rbind(lev_dist_data, data)
#   }
# }
# 
# write.csv(lev_dist_data, "../data/Homology_data/all_lev_dist_data.csv", row.names=FALSE)

lev_dist_data <- read.csv("../data/Homology_data/all_lev_dist_data.csv")
setnames(lev_dist_data, "Homology", "Levenshtein_dist")
```

# Plot just Run 6 of RSWN-lexicase

We can plot the 25%, 50%, and 75% data for a single run just fine:

```{r}
ggplot(subset(lev_dist_data, problem=="replace-space-with-newline" & treatment=="lexicase" & Run==6), 
       aes(x=Generation, y=Levenshtein_dist, group=Quartile, color=factor(Quartile))) + 
  geom_line()
```

# Plot all the RSWN-lexicase runs

If we plot all the data with colors for the quartiles, it's just a muddy mess:

```{r}
ggplot(subset(lev_dist_data, problem=="replace-space-with-newline" & treatment=="lexicase"), 
       aes(x=Generation, y=Levenshtein_dist, group=interaction(Run, Quartile), color=factor(Quartile))) + 
  geom_line(alpha=0.1)
```

If we split it out with factors, we find that they're really not all that different, although 
(as we'd expect) the numbers get higher as you move from 25% to 75%.

```{r}
ggplot(subset(lev_dist_data, problem=="replace-space-with-newline" & treatment=="lexicase"), 
       aes(x=Generation, y=Levenshtein_dist, group=Run)) + 
  geom_line(alpha=0.1) + 
  facet_grid(. ~ Quartile)
```

Reading these plots is also complicated by (again) the fact that some runs end early, and it's hard to see when that's happening.

# Plotting medians for all three treatments

I tried several different plots, but everything tended towards the muddy mess, especially in black and 
white. So I just used facetting throughout.

## Checksum

```{r}
checksum_medians = subset(lev_dist_data, problem=="checksum" & Quartile==0.5)

ggplot(checksum_medians, aes(x=Generation, y=Levenshtein_dist, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

The lexicase Levenshtein distance numbers are definitely quite a lot higher, and remarkabley tightly grouped 
throughout. (Lexicase isn't that tightly bunched in the other three problems.)
IFS and tournament look pretty much the same. These results are pretty similar to the diversity
plots, but don't tell us anything terribly useful about the lack of clustering.

It's interesting that in both IFS and Tourney you can see collapses in Levenshtein distance
(as nearly vertical lines) which probably represent moments where there's very strong convergence.
Those would be interesting moments to try to explore in the Graph DB.

## Negative-to-zero

```{r}
ntz_medians = subset(lev_dist_data, problem=="negative-to-zero" & Quartile==0.5)

ggplot(ntz_medians, aes(x=Generation, y=Levenshtein_dist, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

This one's _really_ interesting because of the clear split between high and low Levenshtein distance
runs for both 
IFS and Tourney, and for what appears to be some quantizing in both of those plots (horizontal lines with
multiple runs having the same median Levenshtein distance). It would be very interesting to see if 
those "high Levenshtein distance" runs correspond to the few successes for IFS and Tourney.

## Replace space with newline

```{r}
rswn_medians = subset(lev_dist_data, problem=="replace-space-with-newline" & Quartile==0.5)

ggplot(rswn_medians, aes(x=Generation, y=Levenshtein_dist, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

Again, both IFS and Tourney seem to split into "high" and "low" Levenshtein distance runs, with the high Levenshtein distance 
runs looking a lot more like lexicase than the low.

## String length backwards

```{r}
slb_medians = subset(lev_dist_data, problem=="string-length-backwards" & Quartile==0.5)

ggplot(slb_medians, aes(x=Generation, y=Levenshtein_dist, group=Run)) +
  geom_line(alpha=0.1) +
  facet_grid(treatment ~ .)
```

This one's arguably the most interesting for lexicase, as it has the most variation through the run. 
Like `checksum` there appear to be some strong vertical drops in all three treatments that presumably 
represent some sort of diversity collapse.

# Trying to better understand the sharp changes when using non-lexicase selection

We know that with lexicase selection there can be _very_ sharp changes in a population's diversity 
because lexicase can select the same individual nearly all the time to be a parent for the next generation.
With tournament selection, however, a single individual is unlikely to be selected terribly often because
the tournament size limits the number of opportunities it has.

So I fished up a tournament selection `negative-to-zero` run that has some sharp changes. This is the plot of the median Levenshtein distance over time, and there are two very strong drops, i.e., two sharp losses of diversity.

```{r}
ggplot(subset(ntz_medians, Run==2 & treatment=="tourney"),
       aes(x=Generation, y=Levenshtein_dist)) +
  geom_line()
```

Let's replot these with limited numbers of generations so we can see how steep the drops really are:

```{r}
ggplot(subset(ntz_medians, Run==2 & treatment=="tourney"),
       aes(x=Generation, y=Levenshtein_dist)) +
  geom_line() +
  coord_cartesian(xlim = c(9, 20))

ggplot(subset(ntz_medians, Run==2 & treatment=="tourney"),
       aes(x=Generation, y=Levenshtein_dist)) +
  geom_line() +
  coord_cartesian(xlim = c(125, 132))
```

There are two pretty substantial drops in median Levenshtein distance, once between generations 11 and 12
(from 0.93 to 0.60), and an even bigger drop later between generations 127 and 128 (from 0.71 to 0.13).

One question is what the other two quartiles look like in this run:
```{r}
ggplot(subset(lev_dist_data, problem=="negative-to-zero" & Run==2 & treatment=="tourney"),
       aes(x=Generation, y=Levenshtein_dist, color=Quartile, group=Quartile)) +
  geom_line()
```

All three are really steep at the end, although the steepness in the first drop seems to lessen as we
move from the 25% quartile through the median to the 75% quartile. 

If we zoom in on the later drop, 
though, we can see that the 25% quartile starts to drop earlier and more strongly than the other two. 
This suggests that there was an accumulation of very similar individuals around some new "discovery"
that started in generation 126 (maybe even 125 -- we don't have information on the bottom of the 
smallest quartile), which gathers momentum through to the large drop ending at generation 128:

```{r}
ggplot(subset(lev_dist_data, problem=="negative-to-zero" & Run==2 & treatment=="tourney"),
       aes(x=Generation, y=Levenshtein_dist, color=Quartile, group=Quartile)) +
  geom_line() +
  coord_cartesian(xlim = c(125, 132))
```

