---
title: "Tom’s Exploring Clustering"
author: "Tom Helmuth"
date: "March 1, 2015"
output: 
  html_document:
    toc: true
---

# Introduction

Let's load some libraries

```{r, message=FALSE}
library('ggplot2')
library('cluster')
library('apcluster')
```

# Test Data

Here's the spread_data from Nic's doc:

```{r}
group_size = 30
spread_data = data.frame(x = c(rnorm(group_size), rnorm(group_size,  0), 
                               rnorm(group_size, 10), rnorm(group_size, 20), 
                               rnorm(group_size, 20)), 
                         y = c(rnorm(group_size), rnorm(group_size, 20), 
                               rnorm(group_size, 10), rnorm(group_size, 20), 
                               rnorm(group_size)))
spread_data$kind = "Spread"

ggplot(spread_data, aes(x=x, y=y, color=kind)) + geom_point()
```

# Clusters At Distances

Let's try to find the number of clusters at different heights in an agnes dendrogram.

```{r}
spread_agnes = agnes(spread_data[,1:2], stand=FALSE, method="average")
sum(spread_agnes$height>10) + 1 #Number of clusters at least averaging 10 distance apart. Have to add 1 since this is counting the number of cluster splits above the specified height.

plot(spread_agnes, which.plots=2)
```

If you use a different linkage, you'll get a different result.

```{r}
spread_agnes_single = agnes(spread_data[,1:2], stand=FALSE, method="single")
sum(spread_agnes_single$height>10) + 1 #Number of clusters at least averaging 10 distance apart. Have to add 1 since this is counting the number of cluster splits above the specified height.

plot(spread_agnes_single, which.plots=2)
```

# Compare to Scaled Data

Let's add the scaled data, at 1/100 original distances.

```{r}
scaled_data <- as.data.frame(scale(spread_data[, 1:2], center=FALSE, scale=c(100, 100)))
scaled_data$kind = "Scaled"

all_data <- rbind(spread_data, scaled_data)
ggplot(all_data, aes(x=x, y=y, color=kind)) + geom_point()
```

Now let's do the same clustering, showing that we get a different number of clusters since the heights will be much smaller.

```{r}
scaled_agnes = agnes(scaled_data[,1:2], stand=FALSE, method="average")
sum(scaled_agnes$height>10) + 1 #Number of clusters at least averaging 10 distance apart. Have to add 1 since this is counting the number of cluster splits above the specified height.

plot(scaled_agnes, which.plots=2)
```
